---
title: "Application to single-cell RNA Seq data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Real data application}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this tutorial, we illustrate how to apply the functions in the `KmeansInference` package to a  collected initially in Chen et al. (2013). 

First we load the package and read in data from GitHub:
```{r}
library(KmeansInference)

library(ggplot2)
require(RCurl)
require(latex2exp)
require(scales)
require(dplyr)
consistent_color <- scales::hue_pal()(8)
githubURL <- "https://github.com/thefactmachine/hex-binning-gis-data/raw/master/popDensity.RData"
load(url(githubURL))
```

### References

Chen T-W, Wardill TJ, Sun Y, Pulver SR, Renninger SL, Baohan A, Schreiter ER, Kerr RA, Orger MB, Jayaraman V, Looger LL, Svoboda K, Kim DS. Ultrasensitive fluorescent proteins for imaging neuronal activity. Nature. 2013;499(7458):295-300. doi:10.1038/nature12354

Chen YT, Jewell SW, Witten DM. (2021) Quantifying uncertainty in spikes estimated from calcium imaging data. [arXiv:2103.0781](https://arxiv.org/abs/2103.07818) [statME].



