---
title: "Application to single-cell RNA-Seq data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Application to single-cell RNA-Seq data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this tutorial, we illustrate how to apply the functions in the `KmeansInference` package to a dataset originally collected by Zheng et al. (2017). 

First we load the package and read in data from GitHub:
```{r}
library(KmeansInference)
library(ggplot2)
require(RCurl)
require(latex2exp)
require(scales)
require(dplyr)
consistent_color <- scales::hue_pal()(8)
githubURL <- "https://github.com/thefactmachine/hex-binning-gis-data/raw/master/popDensity.RData"
load(url(githubURL))
```







### References

Chen YT and Witten DM. (2022+) Selective inference for $k$-means clustering. arXiv preprint. https://arxiv.org/abs/xxxx.xxxxx.

Gao, L. L., Bien, J., and Witten, D. (2020). Selective inference for hierarchical clustering.  arXiv preprint. arXiv:2012.02936.

Zheng, G. X. Y., Terry, J. M., Belgrader, P., Ryvkin, P., Bent, Z. W., Wilson, R., Ziraldo, S. B., Wheeler, T. D., McDermott, G. P., Zhu, J., Gregory, M. T., Shuga, J., Montesclaros, L., Underwood, J. G., Masquelier, D. A., Nishimura, S. Y., Schnall-Levin, M., Wyatt, P. W., Hindson, C. M., Bharadwaj, R., Wong, A., Ness, K. D., Beppu, L. W., Deeg, H. J., McFarland, C., Loeb, K. R., Valente, W. J., Ericson, N. G., Stevens, E. A., Radich, J. P., Mikkelsen, T. S., Hindson, B. J., and Bielas, J. H. (2017), “Massively parallel digital transcriptional profiling of single cells,” Nature Communications, 8, 14049. https://doi.org/10.1038/ncomms14049.



