---
title: "Software tutorials"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Software tutorials}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  warning = FALSE,
  message = FALSE
)
```

This tutorial demonstrates the basic workflow for `KmeansInference`, including
the more-powerful union selective test implemented in
`kmeans_inference_union()`.

```{r}
library(KmeansInference)
library(ggplot2)
```

## Simulate clustered data

We generate \(n=60\) observations in \(q=2\) dimensions from three Gaussian
clusters with common noise standard deviation \(\sigma=1\). The three means form
an equilateral triangle, so clusters 1 and 3 have different population means.

```{r}
set.seed(2022)
n_per <- 20
n <- 3 * n_per
q <- 2
delta <- 5
sig <- 1

true_clusters <- rep(1:3, each = n_per)
mu <- rbind(
  c(delta / 2, 0),
  c(0, sqrt(3) * delta / 2),
  c(-delta / 2, 0)
)
X <- matrix(rnorm(n * q, sd = sig), n, q) + mu[true_clusters, ]

plot_df <- data.frame(
  X1 = X[, 1],
  X2 = X[, 2],
  true_cluster = factor(true_clusters)
)
ggplot(plot_df, aes(X1, X2, colour = true_cluster)) +
  geom_point(size = 2) +
  coord_equal() +
  labs(x = "Feature 1", y = "Feature 2", colour = "True cluster") +
  theme_classic(base_size = 14)
```

## Run k-means

The package uses `kmeans_estimation()` rather than `stats::kmeans()` because the
selective test needs the full Lloyd path, not only the final clustering. The
final cluster labels are the labels used by the inference functions.

```{r}
k <- 3
kmeans_fit <- kmeans_estimation(X, k, iter.max = 20, seed = 2021, verbose = FALSE)
estimated_clusters <- kmeans_fit$final_cluster
table(true_clusters, estimated_clusters)

plot_df$estimated_cluster <- factor(estimated_clusters)
ggplot(plot_df, aes(X1, X2, colour = estimated_cluster)) +
  geom_point(size = 2) +
  coord_equal() +
  labs(x = "Feature 1", y = "Feature 2", colour = "Estimated cluster") +
  theme_classic(base_size = 14)
```

## Original path-conditioned inference

The original selective test conditions on the entire observed Lloyd path. Here
we test the difference in means between estimated clusters 1 and 3.

```{r}
cluster_1 <- 1
cluster_2 <- 3

path_result <- kmeans_inference(
  X, k = k, cluster_1 = cluster_1, cluster_2 = cluster_2,
  sig = sig, iter.max = 20, seed = 2021, verbose = FALSE
)
summary(path_result)
```

The summary reports the norm of the observed difference in cluster means
(`test_stat`), the original selective p-value (`p_selective`), and the naive
Wald p-value (`p_naive`) that ignores the fact that the clusters were selected
from the same data.

## More-powerful union inference

The union test weakens the conditioning event: it preserves the two tested
observed clusters along the perturbation path, but does not require every
intermediate Lloyd assignment to match the observed path. The returned `p_path`
reproduces the original selective p-value, while `p_union` is the more-powerful
selective p-value.

```{r}
union_result <- kmeans_inference_union(
  X, k = k, cluster_1 = cluster_1, cluster_2 = cluster_2,
  sig = sig, iter.max = 20, seed = 2021, verbose = FALSE
)
summary(union_result)
```

In this realization, both selective tests reject the null hypothesis, and the
union p-value is smaller because it conditions on less information. The
`exhaustive` column indicates whether all adjacent Lloyd regions were explored
up to the tail cutoff controlled by `scan_tail_prob`.

## A null split example

Selective inference is most important when k-means splits one population group
into multiple estimated clusters. We rerun k-means with \(K=4\); in this
realization, estimated clusters 2 and 4 are both mostly drawn from the same true
cluster.

```{r}
k_null <- 4
kmeans_fit_null <- kmeans_estimation(
  X, k_null, iter.max = 20, seed = 2021, verbose = FALSE
)
estimated_clusters_null <- kmeans_fit_null$final_cluster
table(true_clusters, estimated_clusters_null)

plot_df$estimated_cluster_k4 <- factor(estimated_clusters_null)
ggplot(plot_df, aes(X1, X2, colour = estimated_cluster_k4)) +
  geom_point(size = 2) +
  coord_equal() +
  labs(x = "Feature 1", y = "Feature 2", colour = "Estimated cluster") +
  theme_classic(base_size = 14)
```

Testing these two split clusters illustrates the danger of ignoring selection:
the naive p-value is tiny, whereas the union selective p-value is moderate.

```{r}
null_result <- kmeans_inference_union(
  X, k = k_null, cluster_1 = 2, cluster_2 = 4,
  sig = sig, iter.max = 20, seed = 2021, verbose = FALSE
)
summary(null_result)
```

If `sig` is omitted, `KmeansInference` uses the robust median-based estimate
from Chen and Witten (2022+). For a general covariance model, set `iso = FALSE`
and pass `SigInv`, the inverse covariance matrix.
